---
title: "Analysis code for Top-down versus bottom-up theories of phonological
           acquisition: A big data approach"
author: "Christina Bergmann, Sho Tsuji, Alejandrina Cristia"
output:
  pdf_document: default
---
  
# Introduction

This document provides the analysis code for the Interspeech 2017 paper "Top-down versus bottom-up theories of phonological acquisition: A big data approach". 





```{r, setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = TRUE)
#ggplot2::theme_set(langcog::theme_mikabr(base_family = "Ubuntu"))
#source("C:/Users/Sho/Documents/GitHub/metalab - Copy/dashboard/global.R", chdir = TRUE)
library(metafor)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(purrr)
library(langcog)
```

```{r FigureTemplate, include = FALSE}
#colors InPhondbp "skyblue" "darkblue"? InWorddbp "indianred"?
#Source for names: https://www.w3schools.com/colors/colors_hex.asp


apatheme=theme_bw()+
  theme(#panel.grid.major=element_blank(),
        #panel.grid.minor=element_blank(),
        panel.border=element_blank(),
        axis.line=element_line(),
        text=element_text(family='Times'),
        legend.position='none')
```


```{r ReadInPhonDB, include = FALSE}
#This file has been extracted from metalab
dbp<-read.table("data/inphon.txt",sep="/")

dbp <- dbp %>%
  mutate(age.C = (mean_age_1-mean(mean_age_1, na.rm=TRUE))/30.42)
```



```{r ReadInWordDB, include = FALSE}
#Get the data based on the most recent spreadsheet, as metalab data is missing some information
source("scripts/calculateES.R")

```



```{r PrepInPhonDB, include = FALSE}

#N infants
dbp$n<-rowSums( cbind (dbp$n_1,dbp$n_2), na.rm=TRUE)
temp<-aggregate(n~same_infant,dbp,mean)
sum(temp$n)

#code moderators
dbp$exposure_phase<-factor(dbp$exposure_phase)
contrasts(dbp$exposure_phase)=contr.sum(4) 
dbp$same_infant<-factor(dbp$same_infant)


#Note where multiple age groups were tested in the same experiment
dbp$levels<-paste(dbp$study_ID,dbp$native_lang,dbp$infant_type,dbp$contrast_sampa)
dbp$subset<-NA
for(eachunique in levels(factor(dbp$levels))){
  subset(dbp, levels==eachunique)->nowdoing
  dbp$subset[dbp$levels==eachunique]<-ifelse(length(levels(factor(unlist(nowdoing[,"mean_age_1"]))))>1,T,F) 
}

dbp$weights_g <-  1/(dbp$g_var_calc)^2
```

```{r PrepInWordDB, include = FALSE}

db <- db %>%
  mutate(age.C = (mean_age_1-mean(mean_age_1, na.rm=TRUE))/30.42)
```

Read-in and preprocessing steps can be found in the R Markdown document. 

# Analyses

## All data

### Vowel acquisition

```{r AnalysisAllInPhonDB}

native.full<-rma.mv(g_calc, g_var_calc, mods=~ age.C + exposure_phase,random = ~ same_infant| short_cite, data=dbp[dbp$test_lang=="native",], weighted=TRUE)
summary(native.full)

nonnative.full<-rma.mv(g_calc, g_var_calc, mods=~ age.C + exposure_phase,random = ~ same_infant | short_cite, data=dbp[dbp$test_lang=="nonnative",], weighted=TRUE)
summary(nonnative.full)
```



```{r PlotAllInPhonDB, echo = FALSE}
#Create models without age centered for plotting purposes
nonnative.full.ageplot<-rma.mv(g_calc, g_var_calc, mods=~ mean_age_1 + exposure_phase,random = ~ same_infant | short_cite, data=dbp[dbp$test_lang=="nonnative",], weighted=TRUE)

native.full.ageplot<-rma.mv(g_calc, g_var_calc, mods=~ mean_age_1 + exposure_phase,random = ~ same_infant| short_cite, data=dbp[dbp$test_lang=="native",], weighted=TRUE)


full = ggplot(dbp, aes(x = mean_age_1, y = g_calc, col = test_lang))
full + geom_point(aes(size = 1/g_var_calc), alpha = .5) + 
  geom_line(y= 0, linetype="dotted") + 
  geom_abline(intercept = native.full.ageplot$b[1], slope = native.full.ageplot$b[2],col="skyblue",size=1.5)+
  geom_abline(intercept = nonnative.full.ageplot$b[1], slope = nonnative.full.ageplot$b[2],col="darkblue",size=1.5)+
  xlab("Age in months") + 
  ylab("Effect size Hedges' g") +
  scale_color_manual(values=c("skyblue","darkblue")) + 
  #scale_x_continuous(0, 17) + #adapt to cover both age ranges, max age in months for InWorddbp = 16.55
  apatheme +
  theme(text = element_text(size=16)) 
```


## Word Segmentation


```{r AnalysisAllInWordDB}
rma_age = rma.mv(g_calc, g_var_calc, mods = age.C,data = db, random =  ~ same_infant | short_cite)

summary(rma_age)
```


```{r PlotAllInWordDB, echo = FALSE}

rma_age_plot = rma.mv(g_calc, g_var_calc, data = db, mod = mean_age_1, random =  ~ same_infant | short_cite) 

p = ggplot(db, aes(mean_age_1, g_calc)) +
  geom_point(aes(size = weights_g), alpha = .5, colour = "indianred") + 
  geom_line(y= 0, linetype="dotted") + 
  geom_abline(intercept = rma_age_plot$b[1], slope = rma_age_plot$b[2], size=1.5, colour = "indianred") +
  xlab("Age in months") + 
  ylab("Effect size Hedges' g") +
  scale_x_continuous(limits = c(5*30.42,17*30.42), 
                     breaks = c(5*30.42, 7.5*30.42, 10*30.42, 12.5*30.42, 15*30.42), 
                     labels = c("5","7.5","10","12.5","15")) + 
  theme_bw() +
  apatheme 

p
  
```



## Only use data testing at least two age groups on same contrast

### Vowels 

Since nonnative model did not converge, dropping same_infant here! (Significant effect of age remains for both model with or without same_infant). 

```{r AnalysisSubsetVowels}

native.sub<-rma.mv(g_calc, g_var_calc, mods=~ age.C + exposure_phase,random = ~ same_infant | short_cite, data=dbp[dbp$test_lang=="native",], weighted=TRUE,subset=c(subset))
summary(native.sub)

nonnative.sub<-rma.mv(g_calc, g_var_calc, mods=~ age.C + exposure_phase,random = ~ 1 | short_cite, data=dbp[dbp$test_lang=="nonnative",], weighted=TRUE,subset=c(subset))
summary(nonnative.sub)
```


```{r PlotSubsetVowels, echo = FALSE}
native.sub.ageplot<-rma.mv(g_calc, g_var_calc, mods=~ mean_age_1 + exposure_phase,random = ~ 1 | short_cite/same_infant, data=dbp[dbp$test_lang=="native",], weighted=TRUE,subset=c(subset))

nonnative.sub.ageplot<-rma.mv(g_calc, g_var_calc, mods=~ mean_age_1 + exposure_phase,random = ~ 1 | short_cite/same_infant, data=dbp[dbp$test_lang=="nonnative",], weighted=TRUE,subset=c(subset))

sub = ggplot(dbp[dbp$subset==T,], aes(x = mean_age_1, y = g_calc, col = test_lang))
sub + geom_point(aes(size = 1/g_var_calc), alpha = .5) + 
  geom_line(y= 0, linetype="dotted") + 
  geom_abline(intercept = native.sub.ageplot$b[1], slope = native.sub.ageplot$b[2],col="skyblue",size=1.5)+
  geom_abline(intercept = nonnative.sub.ageplot$b[1], slope = nonnative.sub.ageplot$b[2],col="darkblue",size=1.5)+
  xlab("Age in months") + 
  ylab("Effect size Hedges' g") +
  scale_color_manual(values=c("skyblue","darkblue")) + 
  scale_x_continuous(limits = c(0,17*30.42), 
                     breaks = c(0*30.42, 2.5*30.42,5*30.42, 7.5*30.42, 10*30.42, 12.5*30.42, 15*30.42), 
                     labels = c("0", "2.5", "5","7.5","10","12.5","15")) + 
  scale_y_continuous(limits=c(-1,2.5)) +
  apatheme +
  theme(text = element_text(size=16))

```

### Word Segmentation


```{r AnalysisSubsetWords}
db_multiple = db %>%
  filter(multiple_age_groups == 1)

rma_age_m = rma.mv(g_calc, g_var_calc, data = db_multiple, mod = age.C, random =  ~ same_infant | short_cite)

summary(rma_age_m) 
```

```{r PlotSubsetInWordDB, echo = FALSE}

rma_age_plot_m = rma.mv(g_calc, g_var_calc, data = db_multiple, mod = mean_age_1, random =  ~ same_infant | short_cite) 

p = ggplot(db, aes(mean_age_1, g_calc)) +
  geom_point(aes(size = weights_g), alpha = .5, colour = "indianred") + 
  geom_line(y= 0, linetype="dotted") + 
  geom_abline(intercept = rma_age_plot_m$b[1], slope = rma_age_plot_m$b[2], size=1.5, colour = "indianred") +
  xlab("Age in months") + 
  ylab("Effect size Hedges' g") +
  scale_x_continuous(limits = c(5*30.42,17*30.42), 
                     breaks = c(5*30.42, 7.5*30.42, 10*30.42, 12.5*30.42, 15*30.42), 
                     labels = c("5","7.5","10","12.5","15")) + 
  theme_bw() +
  apatheme 

p
  
```

### Joint figure

```{r JointFigPrep, include=FALSE}
native = dbp[dbp$subset==T&dbp$test_lang=="native",]
nonnative = dbp[dbp$subset==T&dbp$test_lang=="nonnative",]

db_multiple$Dataset = "WordSeg"
native$Dataset = "Vowels-Native"
nonnative$Dataset = "Vowels-Nonnative"

inworddb = db_multiple %>%
  select(c(mean_age_1, g_calc, weights_g, Dataset))

alldat = full_join(native, nonnative) %>%
  select(c(mean_age_1, g_calc, weights_g, Dataset)) %>%
  full_join(inworddb)

```


```{r PlotJointFigure, echo = FALSE}


sub = ggplot(alldat, aes(x = mean_age_1, y = g_calc, colour = Dataset, linetype = Dataset)) +
  geom_point(aes(size = weights_g), alpha = 1/3,show.legend = FALSE) +
  geom_line(y= 0, linetype="dotted") + 
  geom_abline(intercept = nonnative.sub.ageplot$b[1], slope = nonnative.sub.ageplot$b[2],col="darkblue",size=1.5)+
  geom_abline(intercept = native.sub.ageplot$b[1], slope = native.sub.ageplot$b[2],col="cornflowerblue",size=1.5)+
  geom_abline(intercept = rma_age_plot_m$b[1], slope = rma_age_plot_m$b[2], size=1.5, colour = "indianred") +
  scale_color_manual(values=c("cornflowerblue","darkblue","indianred")) + 
  xlab("Age in months") + 
  ylab("Effect size Hedges' g") +
  scale_x_continuous(limits = c(0,17*30.42), 
                     breaks = c(0*30.42, 2.5*30.42,5*30.42, 7.5*30.42, 10*30.42, 12.5*30.42, 15*30.42,17.5*30.42), 
                     labels = c("0", "2.5", "5","7.5","10","12.5","15","17.5")) + 
  scale_y_continuous(limits=c(-1.1,1.5)) +
  apatheme +
  theme(text = element_text(size=16),
        legend.position = "bottom",
        legend.key = element_blank(),
        legend.background = element_rect(fill = "transparent")
        )

sub

ggsave("figures/AgeEffectSize.jpg", sub)
ggsave("figures/AgeEffectSize.pdf", sub)
```

```{r PrepCI, include=FALSE}

#confidence lines
pr.native<-as.data.frame(print.list.rma(predict(native.sub.ageplot)))
pr.native$age <- dbp$mean_age_1[dbp$test_lang=="native"&dbp$subset==T]

pr.nonnative<-as.data.frame(print.list.rma(predict(nonnative.sub.ageplot)))
pr.nonnative$age <- dbp$mean_age_1[dbp$test_lang=="nonnative"&dbp$subset==T]

pr.words<-as.data.frame(print.list.rma(predict(rma_age_plot_m)))
pr.words$age <- db_multiple$mean_age_1



```


```{r PlotJointFigureCI, echo = FALSE}


sub = ggplot(alldat, aes(x = mean_age_1, y = g_calc)) +
  #geom_point(aes(size = weights_g), alpha = 1/3,show.legend = FALSE) +
  geom_line(y= 0, linetype="dotted") + 
  geom_abline(intercept = nonnative.sub.ageplot$b[1], slope = nonnative.sub.ageplot$b[2],col="darkblue",size=1.5)+
  geom_smooth(data=pr.nonnative,aes(x=age,y=ci.ub), se = FALSE, col="darkblue",size=0.5, method = "lm", formula = y ~ poly(x, 2))+
  geom_smooth(data=pr.nonnative,aes(x=age,y=ci.lb), se = FALSE, col="darkblue",size=0.5, method = "lm", formula = y ~ poly(x, 2))+
  geom_abline(intercept = native.sub.ageplot$b[1], slope = native.sub.ageplot$b[2],col="cornflowerblue",size=1.5)+
  geom_smooth(data=pr.native,aes(x=age,y=ci.ub), se = FALSE, col="cornflowerblue",size=0.5, method = "lm", formula = y ~ poly(x, 2))+
  geom_smooth(data=pr.native,aes(x=age,y=ci.lb), se = FALSE, col="cornflowerblue",size=0.5, method = "lm", formula = y ~ poly(x, 2))+
  geom_abline(intercept = rma_age_plot_m$b[1], slope = rma_age_plot_m$b[2], size=1.5, colour = "indianred") +
  geom_smooth(data=pr.words,aes(x=age,y=ci.ub), se = FALSE, col="indianred",size=0.5)+
  geom_smooth(data=pr.words,aes(x=age,y=ci.lb), se = FALSE, col="indianred",size=0.5)+
  scale_color_manual(values=c("cornflowerblue","darkblue","indianred")) + 
  xlab("Age in months") + 
  ylab("Effect size Hedges' g") +
  scale_x_continuous(limits = c(0,17*30.42), 
                     breaks = c(0*30.42, 2.5*30.42,5*30.42, 7.5*30.42, 10*30.42, 12.5*30.42, 15*30.42,17.5*30.42), 
                     labels = c("0", "2.5", "5","7.5","10","12.5","15","17.5")) + 
  scale_y_continuous(limits=c(-1.1,1.5)) +
  apatheme +
  theme(text = element_text(size=16),
        legend.position = "bottom",
        legend.key = element_blank(),
        legend.background = element_rect(fill = "transparent")
        )

sub

ggsave("figures/AgeEffectSizeCI.jpg", sub)
ggsave("figures/AgeEffectSizeCI.pdf", sub)
# 
# sub = ggplot(db[db$subset==T,], aes(x = mean_age_1, y = g_calc, col = test_lang))
# sub + 
#   scale_color_manual(values=c("skyblue","darkblue","skyblue","darkblue")) + 
#   geom_point(aes(size = 1/g_var_calc), alpha = .5) + 
#   geom_line(y= 0, linetype="dotted") + 
#   geom_abline(intercept = native.sub.ageplot$b[1], slope = native.sub.ageplot$b[2],col="skyblue",size=1.5)+
#   geom_smooth(data=pr.native,aes(x=age,y=ci.lb), error = FALSE)+
#   geom_line(data=pr.native,aes(x=age,y=ci.ub))+
#   #line and ribbon give the same results
#   #geom_ribbon(data=pr.native,aes(x=age,y=pred,ymin=ci.lb,ymax=ci.ub,fill=test_lang,alpha=0.1), stat="identity")+
#   geom_abline(intercept = nonnative.sub.ageplot$b[1], slope = nonnative.sub.ageplot$b[2],col="darkblue",size=1.5)+
#   geom_line(data=pr.nonnative,aes(x=age,y=ci.lb))+
#   geom_line(data=pr.nonnative,aes(x=age,y=ci.ub))+
#   xlab("Age in months") + 
#   ylab("Effect size Hedge's g") +
# 
#   scale_x_continuous(limits = c(0,17*30.42), 
#                      breaks = c(0*30.42, 2.5*30.42,5*30.42, 7.5*30.42, 10*30.42, 12.5*30.42, 15*30.42), 
#                      labels = c("0", "2.5", "5","7.5","10","12.5","15")) + 
#   apatheme +
#   theme(text = element_text(size=16))
```